Synthesis of Rules for Network Intrusion Detectors

### 2019-10-21 paper/paper.tex
pass on technique

\begin{document}
%
\title{Synthesis of Rules for Network Intrusion Detection}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Alcantara\Comment{\inst{1}}\Comment{\orcidID{0000-1111-2222-3333}} \and
Guilherme Padilha\Comment{\inst{1}}\Comment{\orcidID{1111-2222-3333-4444}} \and
Marcelo d'Amorim\Comment{\orcidID{0000-0002-1323-8769}}}
%
\authorrunning{Alcantara et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Network Intrusion Detection Systems (\nids{}) are a popular mechanism
used by system administrators to defend against network attacks. These
systems monitor the network traffic and flag suspicious network
behavior. Signature-based \nids\ do that by checking the network
traffic against a pre-defined set of rules, which can become obsolete
as attackers learn new strategies to circumvent existing defenses.
This paper proposes \tname{}, a technique that automatically
synthesizes \nids\ rules from positive and negative examples (\ie{},
malicious and benign traffic). \tname{} formulates synthesis as an
optimization problem whose candidate solutions are rules that
maximizes the captured the positive traffic and minimizes the captured
negative traffic. \tname{} bootstraps the search with candidate
solutions that use data from the payload of the malicious message and
produces optimal rule candidates on output. We evaluated \tname{} on a
diverse set of attacks. Results indicate that \Fix{...}

@@

\section{Illustrative Example}
\label{sec:suri-metas-coverage}

This section 1)~shows an example rule of Suricata, 2)~briefly explains
the format of Suricata rules, 3) presents an example of a network
attack, and 4) illustrates how \tname{} performs to synthesize the
rule that would capture the attack.

@@

@@

\subsection{\tname\ on SYN Flood Attack}

@@

@@

@@

\tname{} is a search-based approach that uses positive and negative
examples to synthesize signature-based NIDS rules, such as those from
Snort and Suricata. The technique synthesizes rules in two
steps. First, \tname{} produces a potentially overspecified rule that
captures the malicious traffic provided on input. The rule obtained in
this stage can also capture benign traffic. In this step,
\tname\ \emph{maximizes} the number of properties in the rule that
captures the positive traffic. It extracts the tokens from the payload
of the positive traffic (\ie{}, malicious traffic) and add them as
content properties (see Section \Fix{xxx}). Furthermore, it searches
for additional properties satisfied by the positive traffic. Any
subset of properties found in this stage, by construction, captures
the positive traffic. In the second step, \tname{} uses negative
examples to \emph{minimize} the rule, discarding unnecessary
options. The goal of \tname\ is to create a rule that only captures
the positive traffic. The search is optimal (\ie{}, we reach this
goal) if no benign message is captured during the minimization step.

@@

Options can be specific to a given protocol or general and they can
influence or not triggering an
action. Figure~\ref{fig:distribution-rules-protocol} shows the
distributions of these options per protocol and their relevance for
triggering rule actions.

\pgfplotsset{width=5.5cm,compat=1.8}
\begin{figure}[h!]
  \centering
  \begin{tikzpicture}
    \begin{axis}[
        ybar stacked,
        enlargelimits=0.15,
        legend style={at={(0.5,-0.15)},
          anchor=north,legend columns=-1},
        ylabel={\#rules},
        symbolic x coords={gen, http, icmp, udp, tcp},
        xtick=data,
        nodes near coords,
        nodes near coords align={vertical},
      ]
      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
      \legend{relevant, irrelevant}
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:distribution-rules-protocol}Distribution of
    rules per protocol and relevance for triggering action.}
\end{figure}

>>>>

\begin{document}
%
\title{Synthesis of Rules for Network Intrusion Detectors}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Alcantara\Comment{\inst{1}}\Comment{\orcidID{0000-1111-2222-3333}} \and
Guilherme Padilha\Comment{\inst{1}}\Comment{\orcidID{1111-2222-3333-4444}} \and
Marcelo d'Amorim\Comment{\orcidID{0000-0002-1323-8769}}}
%
\authorrunning{Alcantara et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Network Intrusion Detection Systems (\nids{}) are a popular mechanism
used by system administrators to defend against network attacks. These
systems monitor the network traffic and flag suspicious network
behavior. Signature-based \nids\ do that by checking the network
traffic against a pre-defined set of rules, which can become obsolete
as attackers learn new strategies to circumvent existing defenses.
This paper proposes \tname{}, a technique that automatically
synthesizes \nids\ rules from positive and negative examples (\ie{},
malicious and benign traffic). \tname{} formulates synthesis as an
optimization problem whose candidate solutions are rules that
maximizes the captured the positive traffic and minimizes the captured
negative traffic. \tname{} bootstraps the search with candidate
solutions that use data from the payload of the malicious message and
produces optimal rule candidates on output. We evaluated \tname{} on a
diverse set of attacks. Results indicate that \Fix{...}

@@

\section{Background}

This paper focuses on signature-based \nids, Suricata, in
particular. But the principles we use are general to any other
signature-based \nids~(\eg{}, Snort~\cite{snort}).

@@

\Fix{tabulate some important rules that we support}

\section{Illustrative Examples}
\label{sec:suri-metas-coverage}

This section briefly presents examples of attacks and rules that are
synthesized with \tname{} to capture those attacks.

@@

\subsubsection{\tname\ on SYN Flood Attack}

@@

\Fix{repeat this for two other attacks. one non-dos/flood non-content}

@@

\tname{} is a technique that uses positive and negative examples to
synthesize signature-based NIDS rules, such as those from Snort and
Suricata. The goal of \tname\ is to create a rule that \emph{only}
captures the positive traffic, \ie{}, the malicious sequence of
messages.

\vspace{1ex}
\noindent\textbf{Overview.}~\tname{} synthesizes rules in two
steps. First, it produces a potentially overspecified rule that
captures the malicious traffic provided on input. In this step,
\tname\ \emph{maximizes} the number of options in the rule that
captures the positive traffic. It extracts tokens from the malicious
message and use them to define the initial rule. Furthermore, it
searches for additional rule option (\eg, \Fix{xxx}) that are not
found in the message data but are satisfied by the traffic. We say
that this rule overfits the problem as it clearly satisfy the success
criterion of only capturing the positive traffic, however it is unable
to generalize to other attacks of the same kind with different
data. In the second step, \tname{} uses negative examples to
\emph{minimize} the rule, discarding unnecessary options. Note that
any subset of options in the rule obtained in this stage captures the
positive traffic by construction. \Fix{...elaborate...}

\subsection{Step 1: Rule Maximization}
\pgfplotsset{width=5.5cm,compat=1.8}

The goal of the first step is to synthesize a rule with the maximum
number of options that satisfy the malicious traffic. A rule is
maximal if no other option can be added without making it miss the
traffic.

\begin{wrapfigure}[14]{r}{0.5\textwidth}
  \centering
  \vspace{-5ex}  
  \begin{tikzpicture}
    \begin{axis}[
        ybar stacked,
        enlargelimits=0.15,
        legend style={at={(0.5,-0.15)},
          anchor=north,legend columns=-1},
        ylabel={\#rules},
        symbolic x coords={gen, http, icmp, udp, tcp},
        xtick=data,
        nodes near coords,
        nodes near coords align={vertical},
      ]
      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
      \legend{relevant, irrelevant}
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:distribution-rules-protocol}Distribution of
    rules per protocol and relevance for triggering action.}
\end{wrapfigure}
It is important to note that options can be specific to a given
protocol or general and they can influence or not triggering an
action. Figure~\ref{fig:distribution-rules-protocol} shows the
distributions of these options per protocol and their relevance for
triggering rule actions. This plot indicates that the problem search
space is defined based on the protocol of the malicious traffic. For
instance, if \tname{} detects that the attack involves \CodeIn{http}
traffic, then 34 different rule options will be considered in the
creation of the rule.

\Fix{describe extraction...}

\subsection{Step 2: Rule Minimization}

Any rule created from a subset of rule options still captures the
traffic as every rule option we infer in this step is independent and
satisfied by the traffic.

@@

@@
